---
title: "paper2_resistance_mechanisms_mixtures"
output: pdf_document
#output: word_document
author: "Andy South, Ian Hastings and Bethany Levick"
date: "`r Sys.Date()`"
---

```{r,echo=FALSE,results='hide'}
  library(resistance)
```

### version 1
 
# Mechanisms by which insecticide resistance can be influenced by strategies of insectcide use : a simple theoretical exploration. 
  
#### (This file is automagically generated by R from paper2_resistance_mechanisms_mixtures.Rmd)
#### Target : Malaria journal

## Abstract

This paper explores the mechanisms by which insectide resistance is expected to increase with the use of insecticides in isolation, sequence/rotation and mixtures. It is aimed at a readership with limited experience of population genetics and resistance. It uses a model representing two independent loci coding for resistance to two insecticides. It explores factors that are expected to influence the development of resitance, notably the starting frequency of resistance alleles, their dominance, the exposure of insects to the insecticides, the effectiveness of the insecticides, the selective adavantage of resistance and the dominance of the alleles. There is uncertainty about the values of these parameters in field situations. We demonstrate how the values of these parameters are expected to influence the development of resistance and the relative benefits of alternative insecticide application strategies. We describe an R package and user-interface that allows further exploration of the development of resistance.

## Introduction

Insecticide resistance is a problem for malaria (GPIRM), other vector borne diseases and agriculture.

Insecticide resistance becomes a problem when genes coding for resistance firstly arise in a population and secondly increase in frequency. We concern ourselves with this second process of how insecticide resistance increases in frequency within a population.

The changing frequency of insecticide resistance is a population genetic process that can be influenced by :


Parameter  | Description
------------------- | -------------------------------------------------------
1. Effectiveness | proportion of exposed insects with the susceptible allele that are killed by the insecticide
2. Exposure | proportion of the insect population that are exposed to the insecticide
3. Dominance | dominance of the resistance allele relative to the susceptible allele
4. Selection Coefficient  | selective advatange of resistance in the presence of the insecticide
5. Frequency  | frequency of resistance alleles within the population





We use a 2-locus population genetic model of insecticide resistance (Hastings et al. in press) to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies. Hastings et al. (in press) describe the technical details of the model and explore large parameter ranges. In contrast we provide an accessible summary of the model and use selected parameter values to demonstrate the mechanisms by which different parameter values are expected to influence the devlopment of resistance.


## Methods

*accessible summary of the model* 



## Results

* explore the influence of the parameters on sole use either here or in the discussion *


**xA. When all parameters are equal resistance arises slower for sequential than mixture.**
This is because resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time. When used in sequence the curve for each insecticide individually is steeper but, because they happen one after the other, it takes longer to reach the resistance threshold for both.
```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
cat("Fig. 1. All control inputs equal at 0.5")
```

**x1. Increasing the effectiveness of insecticides increases time-to-resistance for mixtures**
A more effective insecticide prompts increased time-to-resistance when used in a mixture. This contrasts with the earlier result showing that when used in isolation a single insecticide prompts a shorter time-to-resistance.

In a mixture the mechanism is that a more effective insecticide provides greater protection to the other. For example taking the previous scenario and increasing the effectiveness of one insecticide from 0.5 to 0.9 increases the time-to-resistance for the 2nd insecticide from around 45 to 75 generations (Fig. 2). In the plot you can see that resistance to insecticide1 (with the increased effectiveness) rises at a similar speed to Fig.2. In contrast resistance to insecticide2 increases more slowly initially. Once insecticide1 reaches it's resistance threshold (around 40 generations), the curve for insecticide2 becomes steeper. This can be explained by the mechanism that once insecticide1 has become ineffective due to resistance it stops killing indivduals that are resistant to insecticide2 and thus stops 'protecting' it from it's own rise in resistance.   

```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
cat("Fig. 2. Effectiveness of insecticide1 increased from 0.5 to 0.9")
```

When the effectiveness of both insecticides are increased to 0.9 the time-to-resistance is increased still further to around 125 generations (Fig. 3). In this case the curves for the rise in resistance to each insecticide are identical. The curves are shallower than previous plots because individuals that are resistant to each insecticide are being killed by the other thus reducing the increase in resistance alleles. As the resistance frequencies go above 10% around generation 100 both insecticides provide less protection to the other and the slope of the resistance curve increases.  


```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.9 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
cat("Fig. 3. Effectiveness of both insecticides 0.9")
```


Increasing the dominance of the alleles within a mixture has the same direction of effect as it does for sole use. Increasing the dominance of the allele corresponding to insecticide1 from 0.5 to 0.9 decreases the time-to-resistance to insecticide1 from 40 to 30 generations but leaves the time-to-resistance for the mixture the same (Fig. 4). Setting the dominance for bothe alleles to 0.9 reduces the time-to-resistance for the mixture to 30 generations (Fig. 5). In both of these cases sequential use continues to have a longer time-to-resistance.

```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
cat("Fig. 4. Dominace of insecticide1 0.9")
```

```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.9 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
cat("Fig. 5. Dominace of both insecticides 0.9")
```



```{r, eval=FALSE, echo=FALSE, fig.height=4}
### 3 examples showing how changing effectiveness can alter the ranking of insecticide strategies
#old modification of curtis values
#### Curtis' Fig 2. inputs : mix2 > adaptive > mix1 > seq
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```

```{r, eval=FALSE, echo=FALSE, fig.height=4}
#old modification of curtis values
#### decrease effectiveness of I2 from 1 to 0.85 : mix2 > adaptive > seq > mix1
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```


```{r, eval=FALSE, echo=FALSE, fig.height=4}
#old modification of curtis values
#### also decrease effectiveness of I1 from 0.73 to 0.63 : seq > mix2 > adaptive > mix1
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.63 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```


## Discussion

**single insecticide**

Parameter to increase  | effect on resistance | Mechanism
------------------ | ------- | ---------------------------------------------------
1. Effectiveness | faster | increased deaths of susceptibles
2. Exposure | faster | increased deaths of susceptibles
3. Dominance | faster | increased survival of heterozygotes
4. Selection Coefficient | faster | increased survival of resistants
5. Frequency  | faster | less change needed to reach resistance threshold

**mixture**

Parameter to increase  | effect on resistance | Mechanism
------------------ | ------- | ---------------------------------------------------
1. Effectiveness | **slower** | increased deaths of individuals resistant to the other insecticide
2. Exposure | **slower** | increased deaths of individuals resistant to the other insecticide
3. Dominance | ?faster | increased survival of resistants
4. Selection Coefficient | ?faster | increased survival of resistants
5. Frequency  | ?faster | less change needed to reach resistance threshold

**difference between mixture and sequential use**

Parameter to increase  | effect on difference | Mechanism
------------------ | ------- | ---------------------------------------------------
1. Effectiveness | improves mixture | faster resistance for single use and slower resistance in mixture
2. Exposure | improves mixture | faster resistance for single use and slower resistance in mixture
3. Dominance | ? | ?
4. Selection Coefficient | ? | ?
5. Frequency  | ? | ?

or like this ?

**Summary of the effect of changing parameters on relative performance of mixture versus sequential strategy**

Parameter  | relative performance mix versus seq | Mechanism
------------------ | ------- | ---------------------------------------------------
x1. High Effectiveness | mix > seq | A highly effective insecticide in a mixture reduces the increase in resistance of the other. 
x2. High Exposure | mix > seq | High exposure is needed for one insecticide in the mixture to reduce the increase of the other.
x3. Dominance | ... | ...
x4. Selection Coefficient | ...
x5. Frequency  | faster | ...



(@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.

(@) For a mixture of 2 insecticides with differing effectiveness. Resistance to the insecticide with a greater effectiveness increases faster. The less effective insecticide is 'protected' by the more effective. Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level.




## References
1. GPIRM
1. Curtis 
1. Mani GS. Evolution of resistance in the presence of two insecticides. Genetics 1985;109(4):761–83.
1. Birget 2015



## Todo

1. create some code to enable plotting 2 scenarios side-by-side
1. re-read GPIRM
1. can I develop explanatory diagrams to describe the input parameters ?
1. mention generation time of 1 month
1. look at effect of sex linkage





