---
title: "paper2_resistance_mechanisms_mixtures"
output: pdf_document
#output: word_document
author: "Andy South, Ian Hastings and Bethany Levick"
date: "`r Sys.Date()`"
---

### version 1
 
# The development of insecticide resistance with use of insecticide mixtures and sequences : a simple theoretical exploration. 
  
#### (This file is automagically generated by R from paper2_resistance_mechanisms_mixtures.Rmd)
#### Target : Malaria journal

## Abstract

This paper explores the mechanisms by which insectide resistance is expected to increase with the use of insecticides in isolation, sequence/rotation and mixtures. It is aimed at a readership with limited experience of population genetics and resistance. It uses a model representing two independent loci coding for resistance to two insecticides. It explores factors that are expected to influence the development of resitance, notably the starting frequency of resistance alleles, their dominance, the exposure of insects to the insecticides, the effectiveness of the insecticides, the selective adavantage of resistance and the dominance of the alleles. There is uncertainty about the values of these parameters in field situations. We demonstrate how the values of these parameters are expected to influence the development of resistance and the relative benefits of alternative insecticide application strategies. We describe an R package and user-interface that allows further exploration of the development of resistance.

## Introduction

Insecticide resistance is a problem for malaria (GPIRM), other vector borne diseases and agriculture.

Insecticide resistance becomes a problem when genes coding for resistance firstly arise in a population and secondly increase in frequency. We concern ourselves with this second process of how insecticide resistance increases in frequency within a population.

The development of insecticide resistance is a population genetic process.

The development of resistance is expected to be influenced by :

1. starting frequency of resistance alleles
1. selection coefficient
   + selective advatange of resistance in the presence of the insecticide
1. dominance
   + dominance of the resistance allele relative to the susceptible allele
1. exposure
   + proportion of the insect population that are exposed to the insecticide
1. effectiveness
   + proportion of exposed susceptible insects that are killed by the insecticide

We use a 2-locus population genetic model of insecticide resistance (Hastings et al. in press) to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies. Hastings et al. (in press) describe the technical details of the model and explore large parameter ranges. In contrast we provide an accessible summary of the model and use selected parameter values to demonstrate the mechanisms by which different parameter values are expected to influence the devlopment of resistance.


## Methods

*Maybe start by exploring the influence of the parameters on sole use ?*


## Results

```{r,echo=FALSE,results='hide'}
  library(resistance)
```
### Some examples showing how changing effectiveness can alter the ranking of insecticide strategies
#### Curtis' Fig 2. inputs : mix2 > adaptive > mix1 > seq
```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```

#### decrease effectiveness of I2 from 1 to 0.85 : mix2 > adaptive > seq > mix1
```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```

#### also decrease effectiveness of I1 from 0.73 to 0.63 : seq > mix2 > adaptive > mix1
```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.63 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```





## Todo

1. look more at GPIRM
1. mention Curtis
1. do I want y axis of resistance frequency plots to be on a log scale ?
1. can I develop explanatory diagrams to describe the input parameters ?
1. decide on base starting scenarios and how to modify params from that
1. mention generation time of 1 month, maybe even rescale axes to years ?





