---
title: "sensiAnPaper1All"
output: html_document
---

### All code for the sensitivity analysis for the first paper

The main question is under which scenarios are mixtures better than sequential ?

This breaks down into these subsidiary quetsions :
1. Under which scenarios is the time to resistance for either insecticide quicker for 

```{r}

  #library(devtools)
  #install_github('AndySouth/resistance')
  library(resistance)

  outFolder <- "C:\\Dropbox\\resistanceResults\\"
```

<!-- eval=FALSE to stop slow code being executed --> 
```{r, eval=FALSE}

  #default setting of the random seed in sensiAnPaperPart() should make the input files mostly the same with exception of exposure
  
  
  #nScenarios <- 100
  nScenarios <- 1000
  ## do model runs
  listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture' )
  listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1' )
  listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2' )
  
  ## save results objects as rda for re-use
  save(listOutMix, file=paste0(outFolder,'listOutMix.rda'))
  save(listOutI1, file=paste0(outFolder,'listOutI1.rda'))
  save(listOutI2, file=paste0(outFolder,'listOutI2.rda'))   
  
    
```  

```{r}
  ## to load previously saved runs, e.g. these that I added 100 onto filename
  load(file=paste0(outFolder,'listOutMix_1000.rda'))
  load(file=paste0(outFolder,'listOutI1_1000.rda'))
  load(file=paste0(outFolder,'listOutI2_1000.rda'))
    
```


### now want to calculate the times to reach critical points and add them on to the input file 
### for different insecticide strategies, sequential, mix1 and mix2
```{r}
  #1) sequential : time to resistance for each insecticide in isolation
  #inputs : inAndOutI1, inAndOutI2
  #find time to criticalPoint for insecticide1
  #find time to criticalPoint for insecticide2
  #add together
  resistPointsI1 <- findResistancePoints(listOutI1, locus=1)
  resistPointsI2 <- findResistancePoints(listOutI2, locus=2)  
  
  resistPointsSeq <- resistPointsI1 + resistPointsI2
  
  
  #2) mixture1 : time to resistance for either insecticide when used in a mixture
  #inputs : inAndOutMix
  #find time to criticalPoint for EITHER insecticide in mixture  
  resistPointsMix1 <- findResistancePoints(listOutMix, locus='either')  
  #todo - to be comparable I think this should be for when resistance to BOTH insecticides is reached
  
  #3) mixture2 : when resistance to one insecticide in the mixture reached, switch to sole use of the 
  #   other until that too reaches the critical point. Record total time.
  # what I actually need to do is start with mixture find the first critical point
  # (need to know which of the insecticides it is)
  # then I need to go to the single run for the other insecticide & starting at 
  # it's current resistance point find out how many more generations to go
  #inputs : inAndOutI1, inAndOutI2, inAndOutMix
  resistPointsMix2 <- findResistancePointsMixResponsive(listOutMix, listOutI1, listOutI2)
  
  #4) mixture3 : time to resistance for both insecticides when used in a mixture
  #inputs : inAndOutMix
  #find time to criticalPoint for BOTH insecticide in mixture  
  resistPointsMix3 <- findResistancePoints(listOutMix, locus='both')
  
```
  
### calculate the difference between sequential & mixture scenarios
```{r}
  
  require('plyr')
  
  ## is mix1 better than sequential ?
  #T/F
  resistBetterMix1SeqBoolean <- resistPointsMix1 > resistPointsSeq
  #convert to 0/1
  resistBetterMix1Seq <- plyr::aaply(resistBetterMix1SeqBoolean,.margins=c(1,2),.fun=as.numeric)
  #rename rows ready for binding on to inputs
  rownames(resistBetterMix1Seq) <- gsub('gen','betterMix1Seq', rownames(resistBetterMix1Seq))
  
  rowSums(resistBetterMix1Seq)/ncol(resistBetterMix1Seq)
  #100runs
  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
  #                .25                   .26                   .30   
  #1000runs
  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
  #               .178                  .232                  .238 
  #new100
  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
  #              0.18                 0.24                 0.25 
  #new1000
  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
  #             0.223                0.245                0.243 

  ## is mix2 better than sequential ?
  #T/F
  resistBetterMix2SeqBoolean <- resistPointsMix2 > resistPointsSeq
  #convert to 0/1
  resistBetterMix2Seq <- plyr::aaply(resistBetterMix2SeqBoolean,.margins=c(1,2),.fun=as.numeric)
  #rename rows ready for binding on to inputs
  rownames(resistBetterMix2Seq) <- gsub('gen','betterMix2Seq', rownames(resistBetterMix2Seq))

    
  rowSums(resistBetterMix2Seq)/ncol(resistBetterMix2Seq)  
  #1000runs
  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
  #             0.471                0.568                0.551 
  #new100
  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
  #              0.58                 0.57                 0.52 
  #new1000
  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
  #             0.558                0.563                0.534   
    
  ## is mix3 better than sequential ?
  #T/F
  resistBetterMix3SeqBoolean <- resistPointsMix3 > resistPointsSeq
  #convert to 0/1
  resistBetterMix3Seq <- plyr::aaply(resistBetterMix3SeqBoolean,.margins=c(1,2),.fun=as.numeric)
  #rename rows ready for binding on to inputs
  rownames(resistBetterMix3Seq) <- gsub('gen','betterMix3Seq', rownames(resistBetterMix3Seq))
  
  #tally runs where mix3 better than seq
  rowSums(resistBetterMix3Seq)/ncol(resistBetterMix3Seq)
  #100runs
  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
  #                  .83                   .75                   .72 
  #1000runs
  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
  #                .794                  .726                  .673  
  #new100
  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
  #              0.83                 0.79                 0.72 
  #new1000
  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
  #             0.820                0.747                0.677   
  
  #check1
  #plot the 2 individual results & the mixture, just to check what is going on
  #x11()
#   plotallele.freq.andy(listOutMix)
#   #x11()
#   plotallele.freq.andy(listOutI1)
#   #x11()
#   plotallele.freq.andy(listOutI2)
  
  #I should be able to compare single scenarios
  #yes this works, but not too useful
  # plotallele.freq.andy( list(results=list(listOutMix$results[[1]], listOutI1$results[[1]] )) )
  #this is a modifdeied version of Beths function
  #allowing curtis fig2 to be applied to any scenario
  #plotcurtis_f2_generic( listOutMix$results[[1]], listOutI2$results[[1]], listOutI1$results[[1]] )
  
  #check2
#   #what is the absolute difference between mix3 & sequential ?
#   resistDifferenceMix3Seq <- resistPointsMix3 - resistPointsSeq
#   
#   #check3 is their a difference between the mix & the individual strategies
#   #looks mostly plausible, occasionally there are 0s
#   resistDifferenceMix1I1 <- resistPointsMix1 - resistPointsI1
#   resistDifferenceMix1I2 <- resistPointsMix1 - resistPointsI2  
  
  ## is mix2 better than sequential ?  
  # I might expect it to be similar to mix3
  
  
  # comparing sensitivity inputs to those used by Beth to reproduce curtis fig2
#   inputCurtis <- createInputMatrix(FALSE)
#   inputSensiMix <- listOutMix$input
#   
#   input <- cbind(inputCurtis, inputSensiMix)
#   write.csv(input, file=paste0(outFolder,'inputCurtisSensiCompare.csv'))
  
  
```

### finding when mixtures are >20% better
### repeating calculation of the difference between sequential & mixture scenarios
```{r}
  
  require('plyr')
  
  ## is mix1 better than sequential ?
  #T/F
  resistBetterMix1SeqBoolean20 <- resistPointsMix1 > (1.2 * resistPointsSeq)
  #convert to 0/1
  resistBetterMix1Seq20 <- plyr::aaply(resistBetterMix1SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
  #rename rows ready for binding on to inputs
  rownames(resistBetterMix1Seq20) <- gsub('gen','betterMix1Seq20', rownames(resistBetterMix1Seq20))
  
  rowSums(resistBetterMix1Seq20)/ncol(resistBetterMix1Seq20)
  #1000runs
 #betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
 #                 0.136                  0.175                  0.173 
  #new1000
 #betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
#                 0.163                  0.178                  0.182   
  
  ## is mix2 better than sequential ?
  #T/F
  resistBetterMix2SeqBoolean20 <- resistPointsMix2 > (1.2 * resistPointsSeq)
  #convert to 0/1
  resistBetterMix2Seq20 <- plyr::aaply(resistBetterMix2SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
  #rename rows ready for binding on to inputs
  rownames(resistBetterMix2Seq20) <- gsub('gen','betterMix2Seq20', rownames(resistBetterMix2Seq20))

    
  rowSums(resistBetterMix2Seq20)/ncol(resistBetterMix2Seq20)  
  #1000runs
#  betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
#                  0.241                  0.326                  0.313 
  #new1000
# betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
#                 0.283                  0.302                  0.283   
    
  ## is mix3 better than sequential ?
  #T/F
  resistBetterMix3SeqBoolean20 <- resistPointsMix3 > (1.2 * resistPointsSeq)
  #convert to 0/1
  resistBetterMix3Seq20 <- plyr::aaply(resistBetterMix3SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
  #rename rows ready for binding on to inputs
  rownames(resistBetterMix3Seq20) <- gsub('gen','betterMix3Seq20', rownames(resistBetterMix3Seq20))
  
  #tally runs where mix3 better than seq
  rowSums(resistBetterMix3Seq20)/ncol(resistBetterMix3Seq20)
  #1000runs
#  betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
#                  0.672                  0.560                  0.488   
  #new1000
# betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
#                 0.622                  0.536                  0.458   
  
```


  
  
### bind results onto input file
```{r}
  
  treeInput <- listOutMix$input
  
  #input files in listOutMix, listOutIn1 & listOutI2 are the same if the runs are done with default randomSeed 
  #except that exposure will be in a.f_AB, a.f_A0 and a.f_B0 respectively (& a.m* too)
  #I could just rename one to exposure
  rownames(treeInput)[rownames(treeInput)=="a.f_AB"] <- "exposure"
  
  #renaming other rownames to make nicer plots
  rownames(treeInput)[rownames(treeInput)=="phi.SS1_A0"] <- "effectiveness_insecticide1"
  rownames(treeInput)[rownames(treeInput)=="phi.SS2_0B"] <- "effectiveness_insecticide2"  

  #add an extra predictor, the lower starting freq of resistance divided by the larger
  startResistanceRatio <- ifelse( treeInput['P_1',] < treeInput['P_2',], treeInput['P_1',]/treeInput['P_2',], treeInput['P_2',]/treeInput['P_1',])
  treeInput <- rbind(treeInput,startResistanceRatio)  
  
  #hardcode which variables to include in the tree analysis here
  #to keep it simple and transparent
  #treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
  #prettifying names
  treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','effectiveness_insecticide1','effectiveness_insecticide2','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
  treeInput <- treeInput[ treePredictors, ]
  
  #replace 0s & 1s with mixture/sequence to make plot clearer
  repMix <- function(x){
    x[x==0] <- 'sequence'
    x[x==1] <- 'mixture'
    x
  }
  
  resistBetterMix1Seq <- repMix(resistBetterMix1Seq)
  resistBetterMix2Seq <- repMix(resistBetterMix2Seq)
  resistBetterMix3Seq <- repMix(resistBetterMix3Seq)

  resistBetterMix1Seq20 <- repMix(resistBetterMix1Seq20)
  resistBetterMix2Seq20 <- repMix(resistBetterMix2Seq20)
  resistBetterMix3Seq20 <- repMix(resistBetterMix3Seq20)

  #bind results onto predictors
  treeInput <- rbind( treeInput, resistBetterMix1Seq )
  treeInput <- rbind( treeInput, resistBetterMix2Seq )
  treeInput <- rbind( treeInput, resistBetterMix3Seq )
  #add >20% results
  treeInput <- rbind( treeInput, resistBetterMix1Seq20 )
  treeInput <- rbind( treeInput, resistBetterMix2Seq20 )
  treeInput <- rbind( treeInput, resistBetterMix3Seq20 )  

```
  
### do tree analysis
```{r, eval=TRUE}
  require(rpart.plot)
  
  #transpose
  treeInput <- t(treeInput)
  
  #convert to a dataframe
  treeInput <- data.frame(treeInput, stringsAsFactors = FALSE)
  
  
  #convert predictor columns back to numeric
  #following this above : replace 0s & 1s with mixture/sequence to make plot clearer
  filter <- !substr(names(treeInput),1,6) %in% 'better'
  treeInput[,filter] <- lapply(treeInput[,filter],as.numeric) 
  filter <- !filter
  treeInput[,filter] <- lapply(treeInput[,filter],as.factor) 
    
  #create string with predictor names in
  treePredictorString <- paste(treePredictors, collapse="+")
  
  
  treeResponses <- c( rownames(resistBetterMix1Seq),
                      rownames(resistBetterMix2Seq),
                      rownames(resistBetterMix3Seq),
                      rownames(resistBetterMix1Seq20),
                      rownames(resistBetterMix2Seq20),
                      rownames(resistBetterMix3Seq20)
                      )
  
  #to do trees & plots for all response variables
  for( treeResponse in treeResponses )
  {
    x11()
    #png(paste0(outFolder,"tree",treeResponse,".png"))
    
    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
    #treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
    
    #http://stackoverflow.com/questions/29197213/what-is-the-difference-between-rel-error-and-x-error-in-a-rpart-decision-tree
    #A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror.
    
    cpOptimal <- tree$cptable[ tree$cptable[,"rel error"] + tree$cptable[,"xstd"]  < tree$cptable[,"xerror"],"CP"][1]
    
    treePruned <- prune(tree, cp=cpOptimal)
    
    
    #to set box colours
    cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")
    
    #adds in no. scenarios correctly and incorrectly classified
    #prp(tree, extra=1, varlen = 10, main=treeResponse)
    #prp(treePruned, extra=1, varlen = 10, main=treeResponse, col=cols, border.col=cols)
    #prp(treePruned, extra=1, varlen = 0, main=treeResponse, col=cols, border.col=cols)
    prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)
            
    #dev.off()
    #savePlot(paste0(outFolder,"tree",treeResponse,".png"), type='png')
    savePlot(paste0(outFolder,"tree",treeResponse,".jpg"), type='jpg')
        
    #trying to limit num levels a different way using maxdepth
    #tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class', maxdepth=3) 
    #adds in no. scenarios correctly and incorrectly classified
    #prp(tree, extra=1, varlen = 10, main=treeResponse)
    #savePlot(paste0(outFolder,"tree",treeResponse,"depth3.jpg"), type='jpg')
    
  }
```  

### testing tree analysis for paper
```{r, eval=TRUE} 

    x11()
    
    treeResponse <- "betterMix2Seq_cP0.5"
  
    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
    
    #to set box colours
    cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")
    
    #adds in no. scenarios correctly and incorrectly classified
    #prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)    
    
    cpOptimal <- tree$cptable[ tree$cptable[,"rel error"] + tree$cptable[,"xstd"]  < tree$cptable[,"xerror"],"CP"][1]
    
    treePruned <- prune(tree, cp=cpOptimal)
    
    prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)        
    
  
    #comparing it with 0/1 gives the same big tree
    treeInput$betterMix2Seq_cP0.5numeric <- as.numeric(treeInput$betterMix2Seq_cP0.5) 
    
    treeResponse <- "betterMix2Seq_cP0.5numeric"
    
    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
    
    #to set box colours
    cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")

    prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)    
        
    
    #testing alternative pruning
    #A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror.
    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
    
  
```  


  
   
```    
#   treeResponse <- 'betterMix1Seq_cP0.5'
#   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
#   
#   #if get error number of rows of matrices must match (see arg 2)
#   #is because all vals are 0
# 
#   plot(tree) #plots tree
#   text(tree) #labels tree
#   #branch predictors 2 phis & P_2  
#     
#   #for mix3 (resistance to both)
#   x11()
#   treeResponse <- 'betterMix3Seq_cP0.5'
#   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class')  
#   plot(tree) #plots tree
#   text(tree) #labels tree  
#   
#   #branch predictors 2 phis & P_1
#   
#   #or from rpart.plot
#   #prp(tree)
#   
#   #adds in no. scenarios correctly and incorrectly classified
#   prp(tree, extra=1, varlen = 10, main=treeResponse)
#   savePlot(paste0(outFolder,"tree",treeResponse,".png"), type='png')
  
  #prp(tree, extra=1,box.col=c("pink","palegreen3")[tree$frame$yval],cex=0.5,yesno.yshift=-2)
  
    #susanas code
    #pruning tree
    #   treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
    #   plot(treePruned) #plots tree
    #   text(treePruned) #labels tree 
    # prp(prunned_tree, extra=1,box.col=c("pink", "palegreen3")[prunned_tree$frame$yval],cex=0.5,yesno.yshift=-2)
```    

