---
  title: "keep on an old insecticide in a mix ?"
  output: pdf_document
  author: "Andy South"
  date: "`r Sys.Date()`"
---
  
## new analysis April 2016 for paper1 : v2
    
Is keeping on an old insecticide in a mix better than just using a new insecticide on its own ?

Short answer : Yes

(part copied from paper1_results_figs.Rmd)

This analysis uses previous sensitivity analysis runs :
1. select out runs where the starting frequency of resistance to I1 > I2
1. I didn't yet restrict to runs where there was a 10 fold difference in starting frequencies, seemed not to be necessary.
  
  
# Notable results
1. it is always better to keep on the old insecticide rather than using the new insecticide on its own.
1. exposure (to both insecticides) and effectiveness of the old insecticide have the main effect on how much better the mixture is over sole use. Exposure has a positive effect, and effectiveness a negative effect on the benefit of the mixture.
1. dominance and selection coefficient for the new insecticide have a lesser effect (also negative) on the benefit of the mixture 
  
  

```{r,echo=FALSE,results='hide'}

library(resistance)
library(ggplot2)

outFolder <- "C:\\Dropbox\\resistanceResults\\"

## to load previously saved runs
#   load(file=paste0(outFolder,'listOutMix_1000.rda'))
#   load(file=paste0(outFolder,'listOutI1_1000.rda'))
#   load(file=paste0(outFolder,'listOutI2_1000.rda'))

## trying with the extended experiment _ex100
experiment <- 'extended'

load(file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
load(file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
load(file=paste0(outFolder,'listOutI2_ex2_10000.rda'))

# load(file=paste0(outFolder,'listOutMix_ex2_1000.rda'))
# load(file=paste0(outFolder,'listOutI1_ex2_1000.rda'))
# load(file=paste0(outFolder,'listOutI2_ex2_1000.rda'))

# very quick test data
# load(file=paste0(outFolder,'listOutMix_ex2_3.rda'))
# load(file=paste0(outFolder,'listOutI1_ex2_3.rda'))
# load(file=paste0(outFolder,'listOutI2_ex2_3.rda'))


```





```{r,echo=FALSE,results='hide'}
### calculate times to reach critical points and add them on to the input file 
### for different insecticide strategies, sequential, mix1 and mix2  


#1) sequential : time to resistance for each insecticide in isolation
#inputs : inAndOutI1, inAndOutI2
#find time to criticalPoint for insecticide1
#find time to criticalPoint for insecticide2
#add together
resistPointsI1 <- findResistancePoints(listOutI1, locus=1)
resistPointsI2 <- findResistancePoints(listOutI2, locus=2)  
resistPointsSeq <- resistPointsI1 + resistPointsI2


#2) mixture1 : time to resistance for either insecticide when used in a mixture
#inputs : inAndOutMix
#find time to criticalPoint for EITHER insecticide in mixture  
resistPointsMix_1 <- findResistancePoints(listOutMix, locus='either')  
#todo - to be comparable I think this should be for when resistance to BOTH insecticides is reached

#3) mixture2 : when resistance to one insecticide in the mixture reached, switch to sole use of the 
#   other until that too reaches the critical point. Record total time.
# what I actually need to do is start with mixture find the first critical point
# (need to know which of the insecticides it is)
# then I need to go to the single run for the other insecticide & starting at 
# it's current resistance point find out how many more generations to go
#inputs : inAndOutI1, inAndOutI2, inAndOutMix
resistPointsMix_A <- findResistancePointsMixResponsive(listOutMix, listOutI1, listOutI2)

#4) mixture3 : time to resistance for both insecticides when used in a mixture
#inputs : inAndOutMix
#find time to criticalPoint for BOTH insecticide in mixture  
resistPointsMix_2 <- findResistancePoints(listOutMix, locus='both')

```



```{r,echo=FALSE,results='hide'}
### chunk copied from part of one in sensiAnPaper1All
### bind results onto input file

treeInput <- listOutMix$input

#input files in listOutMix, listOutIn1 & listOutI2 are the same if the runs are done with default randomSeed 
#except that exposure will be in a.f_AB, a.f_A0 and a.f_B0 respectively (& a.m* too)
#I could just rename one to exposure
#BEWARE risk if future changes
#1/2/16 don't need this now because I've saved exposure from the single original random value
#rownames(treeInput)[rownames(treeInput)=="a.f_AB"] <- "exposure"


#hardcode which variables to include in analysis to keep it simple and transparent
treePredictors <- c('P_1','P_2','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')

#add these for extended analysis
if (experiment=='extended')
  treePredictors <- c(treePredictors,'male_exposure_prop','correct_mix_deploy')

treeInput <- treeInput[ treePredictors, ]

#add an extra predictor, the lower starting freq of resistance divided by the larger
resist_start_lo_div_hi <- ifelse( treeInput['P_1',] < treeInput['P_2',], treeInput['P_1',]/treeInput['P_2',], treeInput['P_2',]/treeInput['P_1',])
treeInput <- rbind(treeInput,resist_start_lo_div_hi)    

#20160122 add test for Ian of resistance1/resistance2
resist_start_1_div_2 <- treeInput['P_1',]/treeInput['P_2',]
treeInput <- rbind(treeInput,resist_start_1_div_2)   

#renaming other rownames to make nicer plots
rownames(treeInput)[rownames(treeInput)=="phi.SS1_A0"] <- "effectiveness_ins1"
rownames(treeInput)[rownames(treeInput)=="phi.SS2_0B"] <- "effectiveness_ins2"

rownames(treeInput)[rownames(treeInput)=="P_1"] <- "start_freq_allele1"
rownames(treeInput)[rownames(treeInput)=="P_2"] <- "start_freq_allele2"

rownames(treeInput)[rownames(treeInput)=="h.RS1_A0"] <- "dominance_allele1"
rownames(treeInput)[rownames(treeInput)=="h.RS2_0B"] <- "dominance_allele2"

rownames(treeInput)[rownames(treeInput)=="s.RR1_A0"] <- "selection_coef_allele1"
rownames(treeInput)[rownames(treeInput)=="s.RR2_0B"] <- "selection_coef_allele2"


#get the inputs here to use in ggplot investigation of model responses to inputs
#used in a later chunk
ggInput <- treeInput

#create a curtisInputs dataframe for use later
curtisInputs <- data.frame("exposure"=0.9,
                           "effectiveness_ins1"=0.73,
                           "effectiveness_ins2"=1,
                           "start_freq_allele1"=0.01,
                           "start_freq_allele2"=0.01,
                           "dominance_allele1"=0.17,
                           "dominance_allele2"=0.0016,
                           "selection_coef_allele1"=0.23,
                           "selection_coef_allele2"=0.43,
                           "male_exposure_prop"=1,
                           "correct_mix_deploy"=1,
                           "resist_start_lo_div_hi"=1,
                           "resist_start_1_div_2"=1)


```  



```{r, eval=TRUE, echo=FALSE, warning=FALSE} 
# getting data into format for ggplot


#uses ggInput calculated in earlier chunk

#first needs to transpose rows to cols
#ggInput_T <- t(ggInput)

#function to transpose and add strategy column with the passed value
#also add on the input columns  
addStrategyColumn <- function(x, value, inputs){
  x <- as.data.frame( t(x) )
  x$strategy <- value
  
  #inputs <- as.numeric(inputs)
  #transpose inputs
  inputs <- as.data.frame( t(inputs), stringsAsFactors=FALSE )
  #cbind onto the outputs
  x <- cbind(inputs,x)
  x
}    

resistPointsI1_T <- addStrategyColumn(resistPointsI1,"insecticide 1",ggInput) 
resistPointsI2_T <- addStrategyColumn(resistPointsI2,"insecticide 2",ggInput)   
resistPointsSeq_T <- addStrategyColumn(resistPointsSeq,"Sequential",ggInput) 
resistPointsMix_1_T <- addStrategyColumn(resistPointsMix_1,"Mix either",ggInput) 
resistPointsMix_A_T <- addStrategyColumn(resistPointsMix_A,"Mix adaptive",ggInput) 
resistPointsMix_2_T <- addStrategyColumn(resistPointsMix_2,"Mix both",ggInput)     

ggInsOuts <- rbind( resistPointsI1_T, resistPointsI2_T, resistPointsSeq_T, 
                    resistPointsMix_1_T, resistPointsMix_A_T, resistPointsMix_2_T)    

#remove runs that didn't reach a resistance threshold (999), even if just for one strategy
#>1000 excludes sequential strategy that had a 999 in
#BEWARE would need to increase 1000 if I increase max generations in the runs
didntReachThresh <- which(ggInsOuts$gen_cP0.5 == 999 | ggInsOuts$gen_cP0.5 > 500 |
                            ggInsOuts$gen_cP0.25 == 999 | ggInsOuts$gen_cP0.25 > 500 |
                            ggInsOuts$gen_cP0.1 == 999 | ggInsOuts$gen_cP0.1 > 500 )

#subset by runs not making threshold
ggInsOuts <- ggInsOuts[-didntReachThresh,]


#prettify output names
names(ggInsOuts)[names(ggInsOuts)=="gen_cP0.5"] <- "time_to_resistance0.5"
names(ggInsOuts)[names(ggInsOuts)=="gen_cP0.25"] <- "time_to_resistance0.25"
names(ggInsOuts)[names(ggInsOuts)=="gen_cP0.1"] <- "time_to_resistance0.1"

#doing; for all inputs
if (experiment=='extended') {
  num_inputs <- 13
} else {
  num_inputs <- 11  
}

#format data in a different way to enable PRCC on difference between sequential & mix2
#resistPointsMix_A_T
#resistPointsSeq_T

#columns to remove from first
indices1 <- which(names(resistPointsMix_A_T) %in% c("gen_cP0.1","gen_cP0.5","strategy"))
#columns to add to 2nd  
indices2 <- which(names(resistPointsMix_A_T) %in% c("gen_cP0.25")) 
#rename column in 2nd
tmp <- resistPointsSeq_T[indices2]
names(tmp) <- "gen_cP0.25seq"

dif_mixA_seq <- cbind(resistPointsMix_A_T[-indices1], tmp)
dif_mix2_seq <- cbind(resistPointsMix_2_T[-indices1], tmp)  

#remove runs that didn't reach a resistance threshold (999), even if just for one strategy
#>1000 excludes sequential strategy that had a 999 in
#BEWARE would need to increase 1000 if I increase max generations in the runs
didntReachThresh <- which( dif_mixA_seq$gen_cP0.25 > 500 | dif_mixA_seq$gen_cP0.25seq > 500 )
#subset by runs not making threshold
dif_mixA_seq <- dif_mixA_seq[-didntReachThresh,]  

dif_mixA_seq["mixA_minus_seq0.25"] <- dif_mixA_seq["gen_cP0.25"] - dif_mixA_seq["gen_cP0.25seq"]

#for dif_mix2_seq
didntReachThresh <- which( dif_mix2_seq$gen_cP0.25 > 500 | dif_mix2_seq$gen_cP0.25seq > 500 )
#subset by runs not making threshold
dif_mix2_seq <- dif_mix2_seq[-didntReachThresh,]  

dif_mix2_seq["mix2_minus_seq0.25"] <- dif_mix2_seq["gen_cP0.25"] - dif_mix2_seq["gen_cP0.25seq"]

```


```{r,echo=FALSE,results='hide'}
### 4/2016 NEW part to analyse what happens when a new insecticide is added to an existing one
### either to replace it or to continue in a mixture

# find the resistance points for I1 & I2 in a mixture 
#(diff than Mix_1 & 2 which is for 1st & 2nd threshold to be reached)
resistPointsMixI1 <- findResistancePoints(listOutMix, locus=1)
resistPointsMixI2 <- findResistancePoints(listOutMix, locus=2)

resistPointsMixI1_T <- addStrategyColumn(resistPointsMixI1,"I1inMix",ggInput) 
resistPointsMixI2_T <- addStrategyColumn(resistPointsMixI2,"I2inMix",ggInput)     

#and use these from above
#resistPointsI1_T
#resistPointsI2_T

#OldPlusNew 
# find whether I1 or I2 has lower starting freq this is the 'new' insecticide
# if I1 < I2 use resistPointsMixI1_T
# if I2 < I1 use resistPointsMixI2_T                                               

#NewOnly
# if I1 < I2 use resistPointsI1_T
# if I2 < I1 use resistPointsI2_T     

# find whether I1 or I2 has lower starting freq this is the 'new' insecticide
# rember the inputs are copied onto the start of all resistPoints objects

indicesI2_lessthan_I1 <- which( resistPointsI1_T$start_freq_allele2 < resistPointsI1_T$start_freq_allele1 )
#this shows about half, I also checked no equals which is good
#length(indicesI2_lessthan_I2)
#[1] 4983
#check whether this works
#check <- resistPointsI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]
#aha! problem I was having is that I1 & I2 inputs are not swapped around
#if I wanted to swap the inputs for I1 & I2 around, would need to do other coefficients etc.
#much potential confusion
#might be better just to restrict to the runs in which I2 < I1
resistPointsNewOnly <- resistPointsI2_T[indicesI2_lessthan_I1, ]
resistPointsOldPlusNew <- resistPointsMixI2_T[indicesI2_lessthan_I1, ]  

# #this bit now replaced by the 2 previous lines
# #NewOnly first make a copy from existing
# resistPointsNewOnly <- resistPointsI1_T
# #replace those values where I2 is the 'new' insecticide with lower resistance
# resistPointsNewOnly[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')] <-
#    resistPointsI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]  
# 
# #OldPlusNew first make a copy from existing
# resistPointsOldPlusNew <- resistPointsMixI1_T
# #replace those values where I2 is the 'new' insecticide with lower resistance
# resistPointsNewOnly[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')] <-
# resistPointsMixI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')] 


resistPointsNewOnly$strategy <- "new only"
resistPointsOldPlusNew$strategy <- "old plus new"

#BE CAREFUL

#format data to enable PRCC on difference between the 2 strategies
#follow previous naming where
#OldPlusNew="gen_cP0.5", NewOnly="gen_cP0.5seq"


#columns to remove from first
indices1 <- which(names(resistPointsOldPlusNew) %in% c("gen_cP0.1","gen_cP0.25","strategy"))
#columns to add to 2nd  
indices2 <- which(names(resistPointsNewOnly) %in% c("gen_cP0.5")) 
#rename column in 2nd
tmp <- resistPointsNewOnly[indices2]
names(tmp) <- "gen_cP0.5seq"

dif_oldnew_newonly <- cbind(resistPointsOldPlusNew[-indices1], tmp)


#remove runs that didn't reach a resistance threshold (999), even if just for one strategy
#>1000 excludes sequential strategy that had a 999 in
#BEWARE would need to increase 1000 if I increase max generations in the runs
didntReachThresh <- which( dif_oldnew_newonly$gen_cP0.5 > 500 | dif_oldnew_newonly$gen_cP0.5seq > 500 )
#subset by runs not making threshold
dif_oldnew_newonly <- dif_oldnew_newonly[-didntReachThresh,]  

dif_oldnew_newonly["mix_minus_sole0.5"] <- dif_oldnew_newonly["gen_cP0.5"] - dif_oldnew_newonly["gen_cP0.5seq"]
# %add_div
dif_oldnew_newonly["mix_divby_sole0.5"] <- dif_oldnew_newonly["gen_cP0.5"] / dif_oldnew_newonly["gen_cP0.5seq"]

#*****************
# to check
# are there any cases where startfreqI1 > startfreqI2
# by my method above i don't think there should be
# i think min should be 1
min(dif_oldnew_newonly$resist_start_1_div_2)
max(dif_oldnew_newonly$resist_start_1_div_2)  

min( dif_oldnew_newonly$start_freq_allele1 - dif_oldnew_newonly$start_freq_allele2)
max( dif_oldnew_newonly$start_freq_allele1 - dif_oldnew_newonly$start_freq_allele2)

#resistPointsNewOnly
min(resistPointsNewOnly$resist_start_1_div_2)
max(resistPointsNewOnly$resist_start_1_div_2) 

# because I now just use runs where startfreq12 < I1
# resist_start_1_div_2 should be the inverse of resist_start_hi_div_lo
# or at least directly correlated
# yes they are all on a single curved line
# plot( resistPointsNewOnly$resist_start_1_div_2 ~ resistPointsNewOnly$resist_start_lo_div_hi)



#****************
#not sure I need to do this because all results already show mixture better
#now I want to subset just those runs where start freq old is > 10 * new 


#proportion of runs in which mix better than sole
num_mix_better <- length(which(dif_oldnew_newonly["mix_minus_sole0.5"] > 0))
num_all <- nrow(dif_oldnew_newonly)
num_mix_better/num_all
#1
#?now mix always seems to be better
#which kind of makes sense, any insecticide better than none

```

\pagebreak

## Fig. xx1 Recreating analysis from Curtis Table 1 (vi)

```{r, eval=TRUE, echo=FALSE, fig.height=4.5} 

runcurtis_f2( P_1 = 0.001,
P_2 = 0.01,
# dominances
h.RS1_A0 =  1,
h.RS2_0B =  1,
# exposures
exposure =  0.9,
# effectivenesses
phi.SS1_A0 =  1,
phi.SS2_0B =  1,
# selection coefficients
s.RR1_A0 =  1,
s.RR2_0B =  1,
addCombinedStrategy=FALSE,
addStrategyLabels = FALSE )
``` 
 
This seems to show that resistance to the new insecticide (red) rises faster when alone (dotted) than when in combination (solid) with an old insecticide at a higher frequency. This is contrary to what Curtis said.

Also note that resistance arises much faster than in our sensitivity runs. I think  this is because the selection coefficients (1) are much higher than the range we use (0.1-0.45).  

\pagebreak

## Fig. xx3 Using a new insecticide on its own or in combination with old

```{r, eval=TRUE, echo=FALSE, fig.height=3.5} 

ggSubset <- dif_oldnew_newonly

names_inputs <- names(ggSubset)[1:num_inputs]


for(i in names_inputs)
{
  #x11()
  
  y <- "mix_minus_sole0.5"    
  
  print( ggplot(ggSubset, aes_string(x=i, y=y)) + #, colour="strategy")) + 
           geom_point(shape=1, colour="darkblue", alpha=0.2, show.legend=FALSE) + 
           #geom_smooth(colour='red', linetype='dashed',size=0.5) +
           geom_smooth(linetype='dashed',size=1, colour="red") +
           #facet_wrap( ~ strategy) +
           #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
           labs(title = i) +
           geom_hline(yintercept = 0, colour="white")
  )
}

#mix_divby_sole0.5
for(i in names_inputs)
{
  #x11()
  
  y <- "mix_divby_sole0.5"    
  
  print( ggplot(ggSubset, aes_string(x=i, y=y)) + #, colour="strategy")) + 
           geom_point(shape=1, colour="darkblue", alpha=0.2, show.legend=FALSE) + 
           #geom_smooth(colour='red', linetype='dashed',size=0.5) +
           geom_smooth(linetype='dashed',size=1, colour="red") +
           #facet_wrap( ~ strategy) +
           #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
           labs(title = i) +
           geom_hline(yintercept = 0, colour="white")
  )
}


```

\pagebreak


# Fig xxx PRCC difference in time-to-resistance between sole new and mix old+new
```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5} 


library(sensitivity)

x <- dif_oldnew_newonly[,1:num_inputs]

y <- dif_oldnew_newonly["mix_minus_sole0.5"]

pcc_res <- pcc(x, y, rank=TRUE)

to_plot <- pcc_res$PRCC
#rename column 1 from 'original to PRCC
names(to_plot)[1] <- 'PRCC'
to_plot$inputs <- rownames(to_plot)  

print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
         geom_point(shape=1, colour='red') +
         theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
         geom_hline(yintercept = 0, linetype=3) +
         ylim(-1,1) +
         ggtitle(paste("Difference time-to-resistance for mix old+new minus new alone (old=1, new=2)")) +
         xlab(NULL)
)

#mix_divby_sole0.5
y <- dif_oldnew_newonly["mix_divby_sole0.5"]

pcc_res <- pcc(x, y, rank=TRUE)

to_plot <- pcc_res$PRCC
#rename column 1 from 'original to PRCC
names(to_plot)[1] <- 'PRCC'
to_plot$inputs <- rownames(to_plot)  

print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
         geom_point(shape=1, colour='red') +
         theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
         geom_hline(yintercept = 0, linetype=3) +
         ylim(-1,1) +
         ggtitle(paste("Ratio of time-to-resistance for mix old+new over new alone (old=1, new=2)")) +
         xlab(NULL)
)



```
