---
title: "Figs for public health journal article April 2017"
output: 
  pdf_document:

#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

```{r, eval=FALSE, include=FALSE}
# options to create final publication quality figures
# causes the document to fail, but the figs do get stored in figs/
library(knitr)
opts_chunk$set(dev="tiff",
               dev.args=list(compression="lzw"),
               dpi=300,
               cache=FALSE,
               fig.path='figs/')
```

```{r,echo=FALSE,results='hide',warning=FALSE}
  library(resistance)
  library(ggplot2)
```

```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="The effect of model inputs on the fitness of genotypes for a single insecticide. Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis. Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed). For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of restoration determines how the fitness for SR lies between that of SS and RR. For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR. In this example effectiveness=0.8, dominance of restoration=0.6, resistance restoration=0.5, dominance of cost=0.6 and resistance cost=0.3."}
  
plot_fit_calc( simple=TRUE, effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.6, dominance_cost=0.6, cost=0.3 )


```

![Screenshot of one online model user interface, accessible at : <https://andysouth.shinyapps.io/resistmob2/>. The user can modify values of the input parameters considered in this paper using simple sliders and run the model to get graphs of resulting resistance frequency over time. Two scenarios (A and B) can be run and the results viewed side by side. This makes it easy to explore the effect of changing individual inputs.](paper2_UI1.png)


#TODO decide which of later figures to modify for pubhealth J paper



```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Comparing time-to-resistance for mixtures and sequences  A. Resistance arises slower for the mixture, B. Resistance arises slower for the sequence. All inputs set to 0.5 except for starting frequency of resistance set to 0.01, and effectiveness of insecticide 1 set to 0.4 in A and 0.8 in B."}

#2 panel fig A. sequence better, B. mix better

#eff1  0.4  0.8

# add side panels for labels
#layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )

layout( matrix(c(1:2,3,3),2,2, byrow = TRUE), heights=c(1,0.15), widths=c(1,1) )
#layout.show(n=3)


#eff1 0.4 seq better

#plot margins, hight t just a bodge to get graphs down closer to legend
par(mar=c(1.5, 4, 15, 0)) #b,l,t,r default c(5, 4, 4, 2)

runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.4 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 140, labelMixSeqRatio = 1 )

mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r

#plot margins, hight t just a bodge to get graphs down closer to legend
par(mar=c(1.5, 4, 15, 0)) #b,l,t,r default c(5, 4, 4, 2)

#eff1 0.8 mix better
runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 140, labelMixSeqRatio = 1 )

mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r


#adding legend in lower panel
#plot margins
par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)

plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m') 

legend( 'bottom', inset=0.1, legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   "), col=c("red","blue","red","blue"), lty=c(2,2,1,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.9, horiz =TRUE )


```
